What can GitHub and Googlecode Issues tell us about Exceptions in Android?

### 2014-08-31 android-stacks.tex
study results

\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
type hierarchy of a Java exception until it reaches a base exception type.

\noindent\emph{Exception Signaler Classifier} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Aplication, Library, Libcore, or Java) 
based on pattern matching between the signaler name and the packages associated to each category.
TELL THAT ALL PACKAGES COMPOSISNG APPS .... WERE RECOVED....

\textbf{Step 4.}  Bytecode-Code analysis of JVM and Android Exceptions. We developerd a bytecode analysis tool based on DesignWizard to discover all exceptions thrown by the last version of SDK version 7, and Android Platform and OS (level 19). Every exception was extracted and classified as Runtime, Checked or Error based on the analysis of the exception hierarchy. Our analysis found XXX runtime exceptions in JVM and XX checked, on Android xxx checked and YYY runtime.

\textbf{Step 5.}  Source-Code Analysis of Android Projects. 
For each one of the analyzed projects, we
downloaded the source code and using custom scripts, we extracted their package
hierarchy (recursive list of of Java packages) and their custom exception types.

\textbf{Step 6.}  Manual Inspection. XXXX of the exceptions reported on stack traces  could not be found on JMV, SO or inside the last version of the project. Based on the full name of the exception, a manual inspection was performed on the artifacts associated to the project and libraries used by the project. When even on those sources the exception was not foung [grep-code and similar sites] the exception, in order to discover the exception type. Only 3 exceptions remained undefined and were removed from this study.

\subsubsection{The ExceptionMiner Tool}

Before investigating the research questions listed above we analyzed the prevalence of stack traces 
on issues created on Github. Prevalence or prevalence proportion, a term often used in epidemiology,
represents the proportion of a population found to have a particular condition
(e.g. a disease) over the total number of the studied population. In our work, we used
this term to evaluate a condition (i.e., presence of a stack trace) over issues
reported on GitHub repositories.

Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined on issues and issue comments. Executing ExceptionMiner tool each issue and issue comment defined on the 16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at least 1 issue on which a stack trace was found. Overall the number of issues on which a stack trace was found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined more than one stack trace.

We then analyzed the lables associated to such issues, and we discovered that only 5,196 were labled as defect ~\footnote{the dedect lables considered in this work were every lable containing as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the issues on which a stack trace was found), most of the issues on which stack traces were found contained no lable. Such observation supports our assumption to consider not only the issues related with defect issues on this analysis.

\subsubsection{The Minig Process}

{MOSTRAR UMA FIGURA COM TODOS OS PASSOS E RESSALTAR OS PASSOS MANUAIS}
COMPLETE HERE.....

\subsubsection{Database of Exception Stack Trace Information}
\subsubsection{Distiling Exception Stack Trace Information}

\- root exception
\- signaler
\- method
\- class
\- package
\- Exception

\subsection{Process}

ESPALHAR ESTA SECAO NAS OUTRAS.

Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
conducted to answer these questions. The mixed-methods approach is composed by the following steps:

@@

\subsection{The Most Common Root Exceptions }

After distilling the information available on the XXXX  stack traces, extracted from the 
issues and issues comments of XXXX Android projects on GitHub and Googlecode,  
we found XXXX  distinct exceptions reported as the root causes of stack traces.
Table \ref{tab:toptenandroid} illustrates the top 10 root couses found in the study ranked by the number of distinct
project repositories on which they were reported.

@@

Table \ref{tab:toptenandroid} illustrate the number of times each exception was reported as 
the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
and the percentages of such in relation to all reported exceptions and analyzed projects respectaly.
The table also presents the whether the method responsible for signaling the exceptions were
part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
 \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.

\begin{table}
  \centering
  \begin{tabular}{rp{29em}}
    \hline
    \bfseries{Signaler} & \bfseries{Description} \\
    \hline
    \bfseries{android} & If the exception is thrown by a method defined in Android Platform or OS, or in a JDK library used by them.\\
    \bfseries{app}     & If the exception is thrown by an application method or in a  DK library used by it.\\
    \bfseries{libcore} & If the exception is thrown by one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
    \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
    \hline
  \end{tabular}
  \caption{Sources of exceptions in Android}
  \label{tab:signalers}
\end{table}

For each exception we could discover the number of times it occurred on exception stack traces,
the number of distinct projects on which such stacks were defined (and consequently its
popularity) and the of times it was signaled by different signaler types. We
could calculate the popularity of each root cause, considering the number of
distinct projects they were found and the whole set of projects analyzed with at
least one stack trace (482). Table~\ref{tab:toptenandroid} 
%and Figure~\ref{fig:androidsignaler}
presents the mined data.

Programming mistakes are the most common root causes for most types of signalers. We
could observe that the NullPointerException is still the exception with higher
number of occurrences. The NullPoiterExceptions are mainly signaled inside
Android platform and inside application code, although we also find
NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw
a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

We could observe in this study that the NullPointerException and other
other implicitly signaled exceptions represented most of the exceptions reported
on issues in both Java and Android subset. For such exceptions, which represent
programming bugs or resource limitations, there is usually no proper handling
besides presenting an error message to the user and restarting the application.
Only high fault tolerant systems need to provide solutions to handle them.

----------

We can observe that most of this set is composed by exceptions defined by java.lang,
which are usually implicitly thrown by JMV dur to a programming mistake. From thi set
the java.lang.NullPointerException  was the most reported root cause;  7578 stack traces 
had a java.lang.NullPointerException  as root cause, which corresponds of 26,31\%  of 
all analyzed stack traces. Considering the number of distinct project repositories, 
the java.lang.NullPointerException, was reported on at least one issue of 1,671 projects 
(which represents 44,47\% of all projects reporting at least one stack trace).

@@

\subsubsection{What They can Tell us?}

@@

INSERIR EM ALGUM PONTO DO TEXTO:
The exceptions available on exception stack traces are not limited to the exceptions 
expliciltly thrown by the Java developer but also include
the exceptions implicitly thrown by the JVM due to programming errors (e.g., out-of-bounds array index, division-by-zero, 
access to a null reference) or resource limitation (e.g., OutOfMemoryError).

@@

TODO: SEPARAR JAVA DE ANDROID

@@

Programming errors are the causes of most of the exception stack traces reported
on issues both in  general Java data set and Android subset. In the Android subset we could find 
more stack traces reporting exceptions related to Backward
compatibility and security concerns than on the general data set.

TODO: Discuss about the interater agreement here...

>>>>

\noindent\emph{Exception Classifier} To support a deeper investigation of the stack traces 
every exception defined on a stack trace need to be classified according to its type
(e.g. Error, Exception or RuntimeException). The exception classifier is based on 
bytecode analysis (based on Design Wizard~\cite{Brunet09}) that walks up the 
type hierarchy of a Java exception until it reaches a base exception type.
This module analyzed the exception hierachy of all exceptions thrown on:
 the last version of SDK version 7, and Android Platform and OS (level 19).

\noindent\emph{Exception Signaler Classifier} 
This module is responsible for classifying each signaler according 
to its origin (i.e. Android Plaform, Android Libcore, Aplication, Library). 
We downloaded the source code of Android Platform and OS (level 19)
and this module extractes the packages hierarchies (recursive list of 
of Java packages), that composes it. To discover the packages that compose 
each application, this module extracted the manifest files of each Android apps
 - which defines the main packages that compose the applications.
Then based on pattern matching between the signaler name and the packages 
 this module identifies the origin of the exception signalers.
The exceptions were considered from libraries if their packages were not defined 
on Android Platform, core libraries nor on applications.
Table ~ref{signalers} summarizes the signaler classification.

\begin{table}
  \centering
  \begin{tabular}{rp{29em}}
    \hline
    \bfseries{Signaler} & \bfseries{Description} \\
    \hline
    \bfseries{android} & If the exception is thrown by a method defined in Android Platform, or in a JDK library used by i.\\
    \bfseries{app}     & If the exception is thrown by an application method or in a JDK library used by it.\\
    \bfseries{libcore} & If the exception is thrown by one of the core libraries reused by Android (i.e., org.apache.harmony, org.w3c.dom, sun.misc, org.apache.http, org.json, org.xml). \\
    \bfseries{lib}     & If the exception is thrown by a method that was not defined by any of the elements above.\\
    \hline
  \end{tabular}
  \caption{Sources of exceptions in Android}
  \label{tab:signalers}
\end{table}

The information extracted by each module was structured in a Database,
which contained the distilled information of stack traces, exception types,
and signalers origin (e.g., root exception and its type, root signaler and its origin,
exception wrapper type, wrapper signaler).

%\subsubsection{The ExceptionMiner Tool}
%Using the ExceptionMiner tool, developed in this work, we could extract the exception stack traces defined 
%on issues and issue comments. Executing ExceptionMiner tool each issue and issue comment defined on the 
%16,836 projects considered in this study, we could observe that 3,758 projects (22,32\%) contained at 
%least 1 issue on which a stack trace was found. Overall the number of issues on which a stack trace was
% found was 21,013 from which 28,800 stack traces were extracted and analyzed - some issues defined 
% more than one stack trace.

%We then analyzed the lables associated to such issues, and we discovered that only 5,196 were
% labled as defect ~\footnote{the dedect lables considered in this work were every lable containing 
%as substring: defect, bug, crash, failure, fail, exception} (which represent only 24,7\% of the
% issues on which a stack trace was found), most of the issues on which stack traces were found 
%contained no lable. Such observation supports our assumption to consider not only the issues 
% related with defect issues on this analysis.

\subsubsection{Manual Inspections}
This study was also based in a set of manual inspection tasks. Besides the 
 identification of Android projects detailes previously. 
Manual inspections were used discover the type of exceptions thrown by 
the applications and libraries we manually inspect the sourcecode or javadoc
 documentation of each of them. When the exception was removed from the 
last app version, the previous versions were analyzed. When the exception was
 not found on the previous versions, the exception was discarded.  Only 3 exceptions 
remained undefined and were removed from this study.

%{MOSTRAR UMA FIGURA COM TODOS OS PASSOS E RESSALTAR OS PASSOS MANUAIS}
%COMPLETE HERE.....
%\subsubsection{Database of Exception Stack Trace Information}
%\subsubsection{Distiling Exception Stack Trace Information}
%\subsection{Process}
%ESPALHAR ESTA SECAO NAS OUTRAS.
%Figure~\ref{fig:overviewfig} presents an overview of the mixed-methods approach
%conducted to answer these questions. The mixed-methods approach is composed by the following steps:

@@

\subsection{The Common Root Exceptions and What They can Tell us? }

After distilling the information available on the exception stack traces we could find 
the the exceptions commonly reported as the root causes of stack traces.
Table \ref{tab:toptenandroid} illustrates the top 10 root couses found in the study,
 ranked by the number of distinct project repositories on which they were reported. 
Thes table shows: the number of times each exception was reported as 
the root cause of a stack trace, the number of distinct projects reporting such causes on stack traces, 
and the percentages of such in relation to all reported exceptions and analyzed projects respectaly.
It also presents the whether the method responsible for signaling the exceptions were
part of Android, the aplication itself, libcore, or a third-party library as presented in Table  \ref{tab:signalers} 
 \footnote{9 stack traces contained only elements of JVM and were discarded from the analysis}.

%For each exception we could discover the number of times it was reported as the root caused on exception stack traces,
%the number of distinct projects on which such stacks were defined (and consequently its
%popularity) and the of times it was signaled by different signaler types. We
%could calculate the popularity of each root cause, considering the number of
%distinct projects they were found and the whole set of projects analyzed with at
%least one stack trace (482). Table~\ref{tab:toptenandroid} 
%and Figure~\ref{fig:androidsignaler} presents the mined data.

@@

INSERIR EM ALGUM PONTO DO TEXTO:

We can observe that most of this set is composed by exceptions defined by java.lang,
which are mostly not expliciltly thrown by the developer but include
the exceptions implicitly thrown by the JVM due to programming errors 
(e.g., out-of-bounds array index, division-by-zero, access to a null reference)
 or resource limitation (e.g., OutOfMemoryError).
For such exceptions, which represent programming bugs or resource limitations, 
there is usually no proper handling besides presenting an error message to
 the user and restarting the application. Only high fault tolerant systems need to 
provide solutions to handle them. Hence, the best strategy to deal with them
is try to prevent them.

\subsubsection{The high number of NullPointerExceptions}
From this set the java.lang.NullPointerException was the most reported root cause; 1225 stack traces 
had a java.lang.NullPointerException  as root cause, which corresponds of 30,08\%  of 
all analyzed stack traces. Considering the number of distinct project repositories, 
the java.lang.NullPointerException, was reported on at least one issue of 99 projects 
(which represents 20,54\% of all projects reporting at least one stack trace).

 The NullPoiterExceptions are mainly signaled inside Android platform and inside 
application code, although we also find
NullPointerExceptions being signaled from third-party libraries. Regarding
reusable code, there is no consensus whether it is a good or bad practice to 
re-throw a NullPointerException. Some prefer to encapsulate such an exception on
IllegalArgumentException, while others~\cite{bloch2008effective} argue that the
NullPointerException makes the cause of the problem explicit and hence 
should not be wrapped.

@@

@@

@@

@@

This classification also enable us to find stack traces reporting exceptions 
directly related to concerns often mentioned on research
works on Android Plaftorm, such as: Concurrency, Backward compatibility, Security, etc.
Considering this broader set the programming mistakesstill represent most of the causes (55,64\%)
of the analyzed exception stack traces.

%TODO: Discuss about the interater agreement here...
